---
title: "Effect of Hyperoxia on development of acute kidney injury in a setting of trauma"
output: html_document
---
---
title: "Hyperoxia in trauma report "
author: "Grigorij Schleifer"
date: "2021"
output:
pdf_document: default
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/grigorijschleifer/Desktop/Hyperoxia-2020/data/")
```

```{r echo=FALSE, warning=FALSE, include=FALSE}
library(tableone)
theme_set(theme_light())
```

```{r echo=FALSE, warning=FALSE, include=FALSE}
# write_csv(cohort_pao2, "~/Desktop/MIMIC_III/PT_Hyperoxia_2020/cohort_pao2.csv")
# write_csv(cohort_pao2_staged, "~/Desktop/MIMIC_III/PT_Hyperoxia_2020/cohort_pao2_staged.csv")
cohort_pao2 <- read_csv("~/Desktop/Hyperoxia-2020/report_data/cohort_pao2.csv")
cohort_pao2_staged <- read_csv("~/Desktop/Hyperoxia-2020/report_data/cohort_pao2_staged.csv")
TableOverall <- read_csv("~/Desktop/Hyperoxia-2020/data/TableOverall.csv")
```


# Explonatory Data Analysis

## AKI variables

aki_stage_48hr <br>
aki_48hr <br>
aki_stage_7day <br>
aki_7day <br>
aki_7day_new

## Inclusion criteria

Age > 18
LOS > 1 and < 7 days

# *Recent* problem

```{r}
# Include only ICU mortality or both? Surviver bias!
# SOFA, OASIS etc as numeric variable for hypothesis testing
```

<br>

Number of patients having a time weighted PaO2 and SpO2 over 100 mmHg and 94 % respectively.

```{r}
## git reset --hard __SHA__
```


```{r echo=FALSE, warning=FALSE}
cohort_pao2 %>%
    nrow()
```

## Table one

```{r echo=FALSE, warning=FALSE}
## numerically coded categorical variables need to be transformed to factors
factorVars <- c(
    "surg", # int
    "has_chronic_lung_conditions", # dbl
    "has_chronic_heart_problems", # dbl
    "has_diabetes", # dbl
    "has_hypertension" # dbl
)

# all variables numeric and categorical
contVars = c(
    "age", # int
    "gender", # factor
    # "insurance",
    "marital_status",
    "ethnicity",
    "religion",
    "admission_type",
    "los", # dbl
    "los_hos", # dbl
    "sofa", # integer
    "saps", # int
    "oasis", # int
    # "age_cat", # factor
    "surg", # int
    "has_chronic_lung_conditions", # dbl
    "has_chronic_heart_problems", # dbl
    "has_diabetes", # dbl
    "has_hypertension", # dbl
    "transfusion_first_24hr", # dbl
    "fluid_intake_first24hr" # dbl
)


# cava NA's for stratification is not possible since NA is dropped
tableOne <-
    CreateTableOne(
        data = cohort_pao2,
        vars = contVars,
        factorVars = factorVars,
        addOverall = FALSE,
        strata = "two_oxy_level"
    )

# chisq.test() for categorical variables (with continuity correction)
# oneway.test() for continous variables (with equal variance assumption, i.e., regular ANOVA) 
# Two-group ANOVA is equivalent of t-test
# if worried about the nonnormal variables and small cell counts
# use the nonnormal argument, as well as the exact (test) argument 
# kruskal.test() is used for the nonnormal continous variables
# fisher.test() is used for categorical variables specified in the exact argument
```


```{r}
tableOne
```

## **Figure 1** AKI proportion based on mortality

```{r echo=FALSE}
cohort_pao2 %>%
    count(icu_mortality, aki_7day_new) %>%
    group_by(icu_mortality) %>%
    mutate(prop = n / sum(n)) %>%
    ggplot(aes(
        fill = as.factor(aki_7day_new),
        y = prop,
        x = as.factor(icu_mortality)
    )) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    theme(legend.position = "right") +
    ylim(c(0, 1)) +
    labs(
        # title = "Proportions in AKI based on ICU mortality",
        # subtitle = "Based on aki_7day_new",
        y = "proportion",
        x = "Died in the ICU",
        # tag = "Figure 1",
        caption = "MIMIC data"
    ) +
    scale_fill_discrete(name = "AKI")
```

## **Figure 2** AKI proportions at increasing _time weighted_ oxygen levels for deceased

```{r echo=F}
cohort_pao2 %>%
    dplyr::filter(hospital_mortality==1) %>% 
    mutate(many_oxy_level = factor(
        many_oxy_level,
        levels = c("<= 75",
                   "> 75 & <= 100",
                   "> 100 & <= 150",
                   "> 150 & <= 200",
                   "> 200")
    )) %>%
    count(many_oxy_level, aki_7day_new, hospital_mortality) %>%
    group_by(many_oxy_level) %>%
    mutate(prop = n / sum(n)) %>%
    ggplot(aes(
        fill = as.factor(aki_7day_new),
        y = prop,
        x = many_oxy_level
    )) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    theme(legend.position = "left") +
    labs(
        # title = "Proportions in AKI based on low vs high oxygen",
        subtitle = "Based on \"aki_7day_new\"",
        y = "proportion",
        x = "High vs low oxygen",
        tag = "Figure 1",
        caption = "MIMIC data"
    ) +
    scale_fill_discrete(name = "AKI")
```

## **Figure 3** AKI proportions at increasing _time weighted_ oxygen levels for survivers

```{r echo=FALSE}
cohort_pao2 %>%
    # dplyr::filter(hospital_mortality==0) %>% 
    mutate(many_oxy_level = factor(
        many_oxy_level,
        levels = c("<= 75",
                   "> 75 & <= 100",
                   "> 100 & <= 150",
                   "> 150 & <= 200",
                   "> 200")
    )) %>%
    count(many_oxy_level, aki_7day_new, hospital_mortality) %>%
    group_by(many_oxy_level) %>%
    mutate(prop = n / sum(n)) %>%
    ggplot(aes(
        fill = as.factor(aki_7day_new),
        y = prop,
        x = many_oxy_level
    )) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    theme(legend.position = "left") +
    labs(
        # title = "Proportions in AKI based on low vs high oxygen",
        subtitle = "Based on \"aki_7day_new\" and all patients",
        y = "proportion",
        x = "High vs low oxygen",
        tag = "Figure 1",
        caption = "MIMIC data"
    ) +
    scale_fill_discrete(name = "AKI")
```

## **Figure 3** AKI proportions at increasing _time weighted_ oxygen levels for ALL patients

```{r echo=FALSE}
cohort_pao2 %>%
    # dplyr::filter(hospital_mortality==0) %>% 
    mutate(many_oxy_level = factor(
        many_oxy_level,
        levels = c("<= 75",
                   "> 75 & <= 100",
                   "> 100 & <= 150",
                   "> 150 & <= 200",
                   "> 200")
    )) %>%
    count(many_oxy_level, aki_7day_new, hospital_mortality) %>%
    group_by(many_oxy_level) %>%
    mutate(prop = n / sum(n)) %>%
    ggplot(aes(
        fill = as.factor(aki_7day_new),
        y = prop,
        x = many_oxy_level
    )) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    theme(legend.position = "left") +
    labs(
        # title = "Proportions in AKI based on low vs high oxygen",
        subtitle = "Based on \"aki_7day_new\" and all patients",
        y = "proportion",
        x = "High vs low oxygen",
        tag = "Figure 1",
        caption = "MIMIC data"
    ) +
    scale_fill_discrete(name = "AKI")
```

Number of patients below/abowe ... time weightednPaO2 value (>300 mmHg)

```{r echo=FALSE}
# aki_7day_new used instead of aki_48h
cohort_pao2 %>% 
    count(hospital_mortality, aki_7day_new) %>% 
    group_by(hospital_mortality) %>% 
    mutate(prop = n/sum(n))
```


**Differences** in maximum 7 day creatinine compared to the admission creatinine in various AKI stages

```{r echo=FALSE}
cohort_pao2_staged %>%
    select(cre_max_7d, admcreat, aki_stage) %>% 
    group_by(aki_stage) %>% 
    summarise(cre_max_7d_mean = mean(cre_max_7d),
              admcreat = mean(admcreat))
```


Time weighted PaO2 distibution over 24 hours

```{r echo=FALSE, warning=FALSE}
cohort_pao2 %>%
    dplyr::filter(!is.na(time_weighted_24hr_pao2), los >= 1) %>%
    ggplot(aes(time_weighted_24hr_pao2)) +
    geom_histogram(binwidth = 10) +
    geom_vline(
        xintercept = median(cohort_pao2$time_weighted_24hr_pao2, na.rm = TRUE),
        color = "green"
    ) +
    geom_vline(
        xintercept = mean(cohort_pao2$time_weighted_24hr_pao2, na.rm = TRUE),
        color = "red"
    ) +
    scale_x_continuous(limit = c(40, 500)) +
    scale_y_continuous(limits = c(0, 200)) +
    xlab("Time-weighted PaO2 (mmHg)") +
    ylab("Count (n)")
```


```{r}
summary(cohort_pao2$time_weighted_24hr_pao2)
```


